Stabilizing selection, archaic introgression, and the distribution of shared functional variation

### 2023-06-27 doc/manuscript.tex
some results

@@

Suppose two pouplation (labeled 0 and 1) diverged some time in the past and
then admix in proportions $f$ and $1-f$.

@@

This shows that additive genetic variance after admixture is equal to the
additive variances in the source populations, weighted by their contributes,
plus a term that depends on the divergence at loci contributing to the trait
between the two populations, measured as $F_2$, weighted by the quadratic
factor $2f(1-f)$.

In general, $F_2$ at a given locus will depend on the demography connecting the
two populations, as well as the effect size at the locus due to selection on
the trait. In the infinitesimal limit, involving many loci each of vanishingly
small effect, dynamics at a given locus will be approximately neutral, so that
$F_2$ will depend only on the demography. In this case
\begin{align*}
    V_G & \approx f V_{G,0} + (1-f) V_{G,1} + 2f(1-f)\mathbb{E}[F_2] \sum_l a_l^2 \\
    & = f V_{G,0} + (1-f) V_{G,1} + 2f(1-f)\mathbb{E}[F_2] n \sigma_m^2
\end{align*}
Here, $\sigma_m^2$ is the variance of effect sizes of new mutations (assuming mean is
zero).

>>>>

\subsection*{Additive genetic variance after admixture}

We expect genetic variance to increase after introgression. The amount that
genetic variance increases depends on the allelic differences accumulated
between populations and the effects of those alleles. Assuming no linkage,
dominance or epistasis, \(V_G=\sum_l 2p_l(1-p_l)a_l^2\). After admixture, with
proportion $f$ contributed by the source population (labeled 0) into the focal
population (labeled 1), \(p_l = fp_{l,0} + (1-f)p_{l,1}\). Plugging into the
expression for $V_G$, and after some simple algebra (Appendix~AX), we can
express the expected genetic variance directly after admixture as \[V_G =
fV_{G,0} + (1-f) V_{G,1} + 2f(1-f)\sum_l F_{2,l} a_l^2,\] where $F_2 = (p_0 -
p_1)^2$ is the squared difference in allele frequencies at a locus
\citep[e.g.,][]{peter2016admixture}. This result in known \citep[e.g.,][]{},
showing that additive variance is equal to that in the source populations
weighted by their contributions, plus a term that depends on the divergence at
trait-affecting loci between the populations weighted by the quadratic factor
$2f(1-f)$.

In general, $F_2$ at a given locus will depend on the demographic history
relating the two populations and the effect size at the locus due to selection
on the trait. In the infinitesimal limit, involving many loci each of
vanishingly small effect, dynamics at a given locus will be approximately
neutral, so that $F_2$ depends only on the demography. In this case
\begin{align*}
    V_G & \approx f V_{G,0} + (1-f) V_{G,1} + 2f(1-f)\mathbb{E}[F_2] \sum_l a_l^2 \\
    & = f V_{G,0} + (1-f) V_{G,1} + 2f(1-f)\mathbb{E}[F_2] n_m V_M,
\end{align*}
where $V_M$ is the variance of effect sizes of new mutations (assuming the mean
is zero).

For $a \not\approx 0$, assuming neutral evolution will tend to overestimate
$F_2$ compared to underdominant selection. We can compute expected $V_G$ both
before and after admixture using the diffusion approximation for the joint SFS
with underdominance. Comparing to simulations with free recombination between
loci (Appendix~AY), we find that this provides an excellent fit to average
observed $V_G$ over time (Figure~\ref{fig:toy-admixture}C,D). Following an
initial rapid increase in $V_G$ due to admixture, it decays back to the
steady-state expectation relatively rapidly.

\begin{itemize}
    \item Benefit of SFS -- partition by frequency classes (prev. seg vs introgressed)
    \item VG from all both previously segregating and introgressed alleles decays,
        replaced by VG due to new mutations
    \item Dividing VG from introgressed alleles into introduced derived vs reintroduced
        ancestral alleles shows the effect of differences in population sizes
\end{itemize}

\subsection*{Recurrent introgression and non-constant population size}

@@

Suppose two population (labeled 0 and 1) diverged some time in the past and
then admix in proportions $f$ and $1-f$.

@@
